---
title: "Acroporid gap analysis"
output:
  word_document: default
  html_document: default
date: "2024-10-05"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = "#>", results='hide', fig.width = 6)
# Load necessary libraries
rm(list=ls(all=TRUE))
library(readxl)
library(dplyr)
library(ggplot2)
library(patchwork)
library(janitor)
library(readr)
library(purrr)
library(fitdistrplus)
library(MASS)
library(systemfonts)
library(flextable)
library(officer)
```

## Exploratory Data Analysis

```{r}
# Load data
raw_ages <- read_excel("raw_ages.xlsx")
raw_mx <- raw_ages |> filter(source == "This paper")

# Select appropriate bandwidth value
bw_rawmx_ucv <- bw.ucv(raw_mx$t_ka)
bw_rawmx_ucv
```

```{r}
# Plot KDE
ggplot(raw_mx) +
  aes(x = t_ka) +
  geom_density(aes(y = after_stat(density), fill = group), alpha = 0.55, adjust = bw_rawmx_ucv) +
  scale_fill_manual(values = c("palmata" = "#dc6601", "Other species" = "gray")) +
  theme(strip.background = element_blank(), panel.background = element_blank(), legend.position = "none") +
  scale_x_continuous(limits = c(0, 6), breaks = seq(0, 6, 0.5)) +
  scale_y_continuous(limits = c(0, 1.5), breaks = seq(0, 1.5, 0.5)) +
  theme(strip.background = element_blank(), panel.background = element_blank(), legend.position = "none", axis.title.x = element_blank()) +
  annotate("rect", xmin = 0.71, xmax = 1.44, ymin = -Inf, ymax = Inf, alpha = 0.1, fill = "#DAC778") +
  annotate("rect", xmin = 1.46, xmax = 1.81, ymin = -Inf, ymax = Inf, alpha = 0.1, fill = "#DAC778") +
  annotate("rect", xmin = 2.47, xmax = 2.88, ymin = -Inf, ymax = Inf, alpha = 0.1, fill = "#DAC778") +
  annotate("rect", xmin = 4.97, xmax = 5.46, ymin = -Inf, ymax = Inf, alpha = 0.1, fill = "#DAC778") +
  annotate("segment", x = 0.6, xend = 1.35, y = 1.25, yend = 1.25, color = "black", linewidth = 2.25) +
  annotate("segment", x = 1.9, xend = 2.2, y = 1.25, yend = 1.25, color = "black", linewidth = 2.25) +
  geom_rug(aes(y = 1.5, colour = group), sides = "t", linewidth = 0.5, alpha = 0.5) +
  scale_colour_manual(values = c("palmata" = "#dc6601", "Other species" = "gray"))
```

### Null Model selection (Monte Carlo simulation)

#### scenario (1) the entire range of ages
```{r}
# Exploring theoretical distributions for Monte Carlo simulation
age <- raw_ages |> filter(source == "This paper")
ages <- age$t_ka

# Fitting multiple distributions
fit_exponential <- fitdist(ages, "exp")
fit_lognormal <- fitdist(ages, "lnorm")
fit_gamma <- fitdist(ages, "gamma")
fit_uniforme <- fitdist(ages, "unif")

# Q-Q plot
par(mfrow = c(1, 1))
qqcomp(
  list(fit_exponential, fit_lognormal, fit_gamma, fit_uniforme),
  legendtext = c("Exponential", "Log Normal", "Gamma", "uniform")
)

# Histogram with fitted density curves
denscomp(
  list(fit_exponential, fit_lognormal, fit_gamma, fit_uniforme),
  legendtext = c("Exponential", "Log Normal", "Gamma", "uniform")
)

# Model selection with AIC and BIC
aic_values <- c(fit_exponential$aic, fit_lognormal$aic, fit_gamma$aic, fit_uniforme$aic)
bic_values <- c(fit_exponential$bic, fit_lognormal$bic, fit_gamma$bic, fit_uniforme$bic)

Model.full <- data.frame(
  Data = "All data",
  Distribucion = c("Exponential", "Log Normal", "Gamma", "Uniform"),
  AIC = aic_values,
  BIC = bic_values
)
```

#### scenario (2) Exclusion of the most recent 500 years

```{r}
# Exclusion of the most recent 500 years
ages500 <- age$t_ka[age$t_ka > 0.5]

# Fitting multiple distributions
fit500_exponential <- fitdist(ages500, "exp")
fit500_lognormal <- fitdist(ages500, "lnorm")
fit500_gamma <- fitdist(ages500, "gamma")
fit500_uniforme <- fitdist(ages500, "unif")

# Q-Q plot
qqcomp(
  list(fit500_exponential, fit500_lognormal, fit500_gamma, fit500_uniforme),
  legendtext = c("Exponential", "Log Normal", "Gamma", "uniform")
)

# Histogram with fitted density curves
denscomp(
  list(fit500_exponential, fit500_lognormal, fit500_gamma, fit500_uniforme),
  legendtext = c("Exponential", "Log Normal", "Gamma", "uniform")
)

# Model selection with AIC and BIC
aic_values500 <- c(fit500_exponential$aic, fit500_lognormal$aic, fit500_gamma$aic, fit500_uniforme$aic)
bic_values500 <- c(fit500_exponential$bic, fit500_lognormal$bic, fit500_gamma$bic, fit500_uniforme$bic)

Model.500 <- data.frame(
  Data = "Exclusion of 500 most recent years",
  Distribucion = c("Exponential", "Log Normal", "Gamma", "Uniform"),
  AIC = aic_values500,
  BIC = bic_values500
)
```

###  Null Model analysis results (Best Fit Selection)

```{r}
models <- bind_rows(Model.full, Model.500)

ft <- flextable(models) %>%
  theme_vanilla() %>%
  autofit() %>%
  set_caption(caption = "Analysis results")

print(ft)
```

### Assessing the contribution of the last 0.5 ka 

```{r}
acrop <-raw_ages|>
  filter(source == "This paper")|>
  filter(species_2 == "BRAN")

age_acrop <- acrop$t_ka
# Calculate the bandwidth using unbiased cross-validation & constructing the KDE
bw_mx_ucv <- bw.ucv(age_acrop)

kde_plot <- ggplot(acrop, aes(x = t_ka)) +
  geom_density(aes(y = ..density..), bw = bw_mx_ucv) +
  labs(title = "Kernel Density Estimate of Age Data",
       x = "Age (ka)",
       y = "Density") +
  theme_minimal()

print(kde_plot)

# Defining the KDE function
kde_function <- density(age_acrop, bw = bw_mx_ucv)

# Calculating the total area under the KDE curve
total_area <- sum(kde_function$y) * diff(kde_function$x)[1]

# Calculating the area for the range 0 to 0.5ka
partial_area <- sum(kde_function$y[kde_function$x >= 0 & kde_function$x <= 0.7]) * diff(kde_function$x)[1]

# Calculate last 0.5ka partial contribution
partial_contribution <- partial_area / total_area

partial_contribution
```

## Monte Carlo Simulation of Gap-size probability

### Simulation using Uniform distribution

## Yucatan

```{r}
age_range <- 1000 * (max(age_acrop) - min(age_acrop))

intervals <- 100 #length of class intervals in years

CI <- round(age_range/intervals, 0) # number of classes

# Histogram to visualize the data
hist(age_acrop,
     breaks = CI, 
     main = paste("Yucatan,", intervals, "years intervals"),
     xlab = "Age (ka)",
     ylab = "Frequency")

# Defining class intervals
breaks <- hist(age_acrop, 
               breaks = CI, 
               main = paste("Yucatan,", intervals, "years intervals"),
               xlab = "Age (ka)",
               ylab = "Frequency")$breaks

# Calculate the observed frequencies
observed <- hist(age_acrop, breaks = breaks, plot = FALSE)$counts

# Function to identify sequences of zeros and their length
gaps <- function(vector) {
  secuencias <- data.frame(inicio = integer(), longitud = integer())
  longitud_actual <- 0
  inicio_secuencia <- NULL
  
  for (i in 1:length(vector)) {
    if (vector[i] == 0) {
      if (is.null(inicio_secuencia)) {
        inicio_secuencia <- i
      }
      longitud_actual <- longitud_actual + 1
    } else {
      if (!is.null(inicio_secuencia)) {
        secuencias <- rbind(secuencias, data.frame(inicio = inicio_secuencia, longitud = longitud_actual))
        longitud_actual <- 0
        inicio_secuencia <- NULL
      }
    }
  }
  
  if (!is.null(inicio_secuencia)) {
    secuencias <- rbind(secuencias, data.frame(inicio = inicio_secuencia, longitud = longitud_actual))
  }
  
  secuencias$size <- secuencias$longitud * (age_range/length(observed))
  
  
  return(secuencias)
}

# Monte Carlo Simulation
n_sim <- 10000
simulated_gaps <- numeric(n_sim)
simulated_gaps_size <- vector(mode = "list", length = n_sim)

for (i in 1:n_sim) {
  # Parameterized Uniform Distribution
  simulated_data <- runif(length(age_acrop), min = min(age_acrop), max = max(age_acrop))
  
  # Calculate the simulated frequencies
  sim_observed <- hist(simulated_data, breaks = CI, plot = FALSE)$counts
  
  # Gap tests
  simulated_gaps[i] <- sum(sim_observed == 0)
  simulated_gaps_size[[i]] <- gaps(sim_observed)
}

# Gap size analysis
gap_size <- list_rbind(simulated_gaps_size, names_to = "simulation")|>
  group_by(simulation, size)|>
  reframe(f = n())

yucatan <- gap_size|>
  group_by(size)|>
  summarise(average = round(mean(f),0), times = length(size), freq = (times+1)/(n_sim+1))|>
  mutate(site = "Yucatan")

# plotting gaps probability in response to gap size

ggplot(yucatan, aes(x = size, y = freq)) +
  scale_y_continuous(breaks = seq(0, 1, 0.2)) +
  scale_x_continuous(breaks = seq(0, 1400, 100)) +
  geom_hline(yintercept = 0.05,
             colour = "black",
             linewidth = 1, 
             linetype = "longdash") +
  ylab(expression("Probability under " * H[0] * " of uniform distribution")) +
  xlab("Gap size (years)") +
  geom_line(aes(color = "N_Yucatan, n=64"), linewidth = 1) +  
  scale_color_manual(values = c("N_Yucatan, n=64" = "darkorange"), name = "Legend") +
  theme_bw() +
  theme(
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank(),
    panel.grid.minor.y = element_blank(),
    panel.grid.major.y = element_blank(),
    legend.position = "none") +
  guides(color = guide_legend(override.aes = list(linetype = 1, size = 1))) +
  annotation_custom(
    grob = grid::grobTree(
      grid::textGrob("N_Yucatan, n=64", x = 0.95, y = 0.95, hjust = 1, gp = grid::gpar(col = "darkorange"))
    ),
    xmin = -Inf, xmax = Inf, ymin = -Inf, ymax = Inf
  )
```

## Gap uncertainties & sample size effect 

### Gap uncertainties analysis

```{r}
#calculating density estimates for the apal ages
pdf0<-density(age_acrop,bw=0.09,from=-0.1,to=6)
#ploting KDE 
plot(pdf0,xlim=c(-0.1,6),ylim=c(0,1.2))
# randomly sampling test
pdf25_1<-density(age_acrop[sample(1:51,25)],bw=0.09,from=-0.1,to=6)
b<-pdf0$x[2]-pdf0$x[1]
sum(abs(pdf0$y-pdf25_1$y)*b)/2

# Creating a function to generate densities with different sample sizes 
generar_densidades <- function(datos, inicio = 36, fin = 64) {
# Creating empty list to store results
  lista_densidades <- list()
# Loop for each sample size
  for(n in seq(inicio, fin, by = 4)) {
# generating Sample 
    muestra <- datos[sample(1:length(datos), n)]
# Calculating density
    densidad <- density(muestra, bw = 0.09, from = -0.1, to = 6)
# Storing densities 
    lista_densidades[[paste0("pdf", n)]] <- densidad
  }
# return densities list
  return(lista_densidades)
}
# function use
densidades <- generar_densidades(age_acrop)
# empty dataframe
df_densidades <- data.frame()
# density loop
for(i in names(densidades)) {
# temporal dataframe
  df_temp <- data.frame(x = densidades[[i]]$x, y = densidades[[i]]$y, group = i)
# adding temporal to main dataframe
  df_densidades <- rbind(df_densidades, df_temp)
}

#Evaluating uncertainties in densities
# Empty dataframe to store density differences
df_diferencias <- data.frame()

# Creating a loop for each density on the list
for(i in names(densidades)) {
  # Creating temporal dataframe to store delta densities
  df_temp <- data.frame(x = pdf0$x, y = pdf0$y - densidades[[i]]$y, group = i)
  
  # Adding temporal dataframe with delta densities to the main dataframe
  df_diferencias <- rbind(df_diferencias, df_temp)
}

#ploting gap uncertainties
library(paletteer)
colores2<-paletteer_c("ggthemes::Orange", 30)
ggplot(df_diferencias, aes(x = x, y = y, fill = group)) +
  annotate("rect", xmin = 0.71, xmax = 1.44, ymin = -Inf, ymax = Inf, alpha = 0.15, fill = "#DAC778") + 
  annotate("rect", xmin = 1.46, xmax = 1.81, ymin = -Inf, ymax = Inf, alpha = 0.15, fill = "#DAC778") +
  annotate("rect", xmin = 2.47, xmax = 2.88, ymin = -Inf, ymax = Inf, alpha = 0.15, fill = "#DAC778") +
  annotate("rect", xmin = 4.97, xmax = 5.46, ymin = -Inf, ymax = Inf, alpha = 0.15, fill = "#DAC778")+
  geom_area(alpha = 0.5) +
  scale_fill_manual(values = colores2) +
  labs(x = "age(ka)", y = "Density uncertainties", fill = "Group") +
  coord_cartesian(xlim = c(-0.1, 6))+
  theme_bw()+
  theme(legend.position = "none") 
```

### Sample size analysis

```{r}
#Accessing resampling related uncertainties
desajuste<-matrix(ncol=1000,nrow=55)
for(j in 1:nrow(desajuste)){
  for(i in 1:ncol(desajuste)){
    desajuste[j,i]<-sum(abs(pdf0$y - 
                              density(age_acrop[sample(1:64,(j+9))],bw=0.09,from=-0.1,to=6)$y)*b)/2
  }
}

dim(desajuste)
# Create a new data frame for plotting PUMA age densities
df <- data.frame(
  x = 1:55,
  mean = apply(desajuste, 1, mean),
  q25 = apply(desajuste, 1, quantile, 0.25),
  q75 = apply(desajuste, 1, quantile, 0.75),
  q025 = apply(desajuste, 1, quantile, 0.025),
  q975 = apply(desajuste, 1, quantile, 0.975)
)

# plotting sample size impact 
ggplot(df, aes(x = x, y = mean)) +
  geom_ribbon(aes(ymin = q025, ymax = q975), alpha = 0.1) +
  geom_ribbon(aes(ymin = q25, ymax = q75), alpha = 0.4) +
  geom_point(shape = 16) +
  labs(x = "resamples", y = "mean") +
  scale_x_continuous(breaks = seq(0, 60, by = 10)) +
  scale_y_continuous(limits = c(0, 0.55)) +
  theme_bw()+
  geom_hline(yintercept = 0.1, color = "black", linetype = "dashed")+
  annotate("rect", xmin = 37, xmax = 55, ymin = -Inf, ymax = Inf, alpha = 0.15, fill = "lightgreen")

```

## Regional trends (Exploratory analysis of published data for comparative purpouses)

```{r}
regional<-read_excel("raw_ages.xlsx")# ADJUSTED TEMPORAL FRAME 

# regional figure
fNW_Car <- regional[regional$spot == "Northwestern Caribbean" , ] # This paper, Mexico
fSW_Car <- regional[regional$spot == "Southwestern Caribbean" , ] # Belize
fE_Car <- regional[regional$spot == "Eastern Caribbean" , ] # Saint Croix
fLA_Car <- regional[regional$spot == "Lesser Antilles"& regional$source =="Abdul et al.,16", ] # Barbados


summary(as.factor(fNW_Car$group))
summary(as.factor(fSW_Car$group))
summary(as.factor(fE_Car$group))
summary(as.factor(fLA_Car$group))

A<-ggplot(fNW_Car) +
  aes(x = t_ka) +
  geom_density(aes(y = after_stat(density), fill = group), alpha = 0.55, adjust = 0.09) +
  scale_fill_manual(values = c("palmata" = "#dc6601","Other species" = "gray")) +
  theme(strip.background = element_blank(), panel.background = element_blank(), legend.position = "none") +
  scale_x_continuous(limits = c(0, 14), breaks = seq(0, 14, 2))+
  scale_y_continuous(limits = c(0, 1.5), breaks = seq(0, 1.5, 1/2))+
  theme(strip.background = element_blank(), panel.background = element_blank(),
        legend.position = "none", axis.title.x = element_blank())+
  annotate("rect", xmin = 0.71, xmax = 1.44, ymin = -Inf, ymax = Inf, alpha = 0.1, fill = "#ee9f27") + 
  annotate("rect", xmin = 1.46, xmax = 1.81, ymin = -Inf, ymax = Inf, alpha = 0.1, fill = "#ee9f27") +
  annotate("rect", xmin = 2.47, xmax = 2.88, ymin = -Inf, ymax = Inf, alpha = 0.1, fill = "#ee9f27") +
  annotate("rect", xmin = 4.97, xmax = 5.46, ymin = -Inf, ymax = Inf, alpha = 0.1, fill = "#ee9f27")+
  annotate("segment", x = 0.6, xend = 1.35, y = 1.25, yend = 1.25, color = "black", linewidth = 2.25) +
  annotate("segment", x = 1.9, xend = 2.2, y = 1.25, yend = 1.25, color = "black", linewidth = 2.25) +
  annotate("text", x = 13.5, y = 1.25, label = "N_Yucatan", color="black")+
  geom_rug(aes(y = 1.5, colour = group), sides = "t", linewidth = 0.5, alpha = 0.5) +
  scale_colour_manual(values = c("palmata" = "#dc6601", "Other species" = "gray"))

B<-ggplot(fSW_Car) +
  aes(x = t_ka) +
  geom_density(aes(y = after_stat(density), fill = group), alpha = 0.55, adjust = 0.088) +
  scale_fill_manual(values = c("palmata" = "#dc6601","Other species" = "gray")) +
  theme(strip.background = element_blank(), panel.background = element_blank(), legend.position = "none") +
  scale_x_continuous(limits = c(0, 14), breaks = seq(0, 14, 2))+
  scale_y_continuous(limits = c(0, 1.5), breaks = seq(0, 1.5, 1/2))+
  theme(strip.background = element_blank(), panel.background = element_blank(),
        legend.position = "none", axis.title.x = element_blank())+
  annotate("rect", xmin = 1.9, xmax = 2.72, ymin = -Inf, ymax = Inf, alpha = 0.1, fill = "#ee9f27") + 
  annotate("rect", xmin = 3.68, xmax = 4.2, ymin = -Inf, ymax = Inf, alpha = 0.1, fill = "#ee9f27") +
  annotate("rect", xmin = 5.54, xmax = 6.0, ymin = -Inf, ymax = Inf, alpha = 0.1, fill = "#ee9f27") +
  annotate("text", x = 13.5, y = 1.25, label = "Belize", color="black") +
  geom_rug(aes(y = 1.5, colour = group), sides = "t", linewidth = 0.5, alpha = 0.5) +
  scale_colour_manual(values = c("palmata" = "#dc6601", "Other species" = "gray"))

C<-ggplot(fE_Car) +
  aes(x = t_ka) +
  geom_density(aes(y = after_stat(density), fill = group), alpha = 0.55, adjust = 0.088, linetype = "solid") +
  scale_fill_manual(values = c("palmata" = "#dc6601","Other species" = "gray")) +
  theme(strip.background = element_blank(), panel.background = element_blank(), axis.title.x = element_blank(),legend.position = "none") +
  scale_x_continuous(limits = c(0, 14), breaks = seq(0, 14, 2))+
  scale_y_continuous(limits = c(0, 1.5), breaks = seq(0, 1.5, 1/2))+
  annotate("rect", xmin = 2.0, xmax = 3.0, ymin = -Inf, ymax = Inf, alpha = 0.1, fill = "#ee9f27") +
  annotate("rect", xmin = 5.2, xmax = 5.9, ymin = -Inf, ymax = Inf, alpha = 0.01, fill = "#ee9f27")+
  annotate("segment", x = 3.6, xend = 4.8, y = 1.25,  yend = 1.25, color = "black",linewidth= 2.25)+
  annotate("segment", x = 1.0, xend = 2.5, y = 1.25,  yend = 1.25, color = "black",linewidth= 2.25)+
  annotate("segment", x = 0.0, xend = 0.3, y = 1.25,  yend = 1.25, color = "black",linewidth= 2.25)+
  annotate("text", x = 13.5, y = 1.25, label = "Saint Croix", color="black")+
  geom_rug(aes(y = 1.5, colour = group), sides = "t", linewidth = 0.5, alpha = 0.5) +
  scale_colour_manual(values = c("palmata" = "#dc6601", "Other species" = "gray"))

D<-ggplot(fLA_Car) +
  aes(x = t_ka) +
  geom_density(aes(y = after_stat(density), fill = group), alpha = 0.55, adjust = 0.088) +
  scale_fill_manual(values = c("palmata" = "#dc6601","Other species" = "gray")) +
  theme(strip.background = element_blank(), panel.background = element_blank(), legend.position = "none") +
  scale_x_continuous(limits = c(0,14), breaks = seq(0, 14, 2))+
  scale_y_continuous(limits = c(0, 1.5), breaks = seq(0, 1.5, 1/2))+
  annotate("rect", xmin = 9.92, xmax = 10.28, ymin = -Inf, ymax = Inf, alpha = 0.1, fill = "#ee9f27") +  
  annotate("text", x = 13.5, y = 1.25, label = "Barbados", color="black")+
  geom_rug(aes(y = 1.5, colour = group), sides = "t", linewidth = 0.5, alpha = 0.5) +
  scale_colour_manual(values = c("palmata" = "#dc6601", "Other species" = "gray"))

regional_palmata <- subset(regional, group == "palmata")

E<-ggplot(regional_palmata, aes(x = t_ka)) +
  geom_density(aes(y = after_stat(density), fill = source), alpha = 0.85, adjust = 0.088) +
  scale_fill_manual(values = c("This paper" = "#dc6601", "Gischler et al.,23" = "#ee9f27","Hubbard et al.,05;13" = "#FCA360", "Abdul et al.,16" = "#FFC125")) +
  theme(strip.background = element_blank(), panel.background = element_blank(), , legend.position = "none") +
  labs(x = "t_ka", y = "Density (after stats)") +
  scale_x_continuous(limits = c(0,14), breaks = seq(0, 14, 2))+
  scale_y_continuous(limits = c(0, 1.5), breaks = seq(0, 1.5, 1/2))+
  geom_rug(aes(y = 1.5, colour = source), sides = "t", linewidth = 0.5, alpha = 0.5) +
  scale_colour_manual(values = c("This paper" = "#dc6601", "Gischler et al.,23" = "#ee9f27","Hubbard et al.,05;13" = "#FCA360", "Abdul et al.,16" = "#FFC125"))

combined_plot <- A/D/B/C/E
combined_plot 
```

## Monte Carlo Simulation of Gap-size probability (Regional Scope).

## Barbados

```{r}

acrop <-raw_ages|>
  filter(source == "Abdul et al.,16")|>
  filter(species_2 == "BRAN")

age_acrop <- acrop$t_ka

age_range <- 1000 * (max(age_acrop) - min(age_acrop))

intervals <- 100

CI <- round(age_range/intervals, 0)

# Histogram to visualize the data
hist(age_acrop,
     breaks = CI, 
     main = paste("Barbados,", intervals, "years intervals"),
     xlab = "Age (ka)",
     ylab = "Frequency")

# Defining class intervals
breaks <- hist(age_acrop, 
               breaks = CI, 
               main = paste("Barbados,", intervals, "years intervals"),
               xlab = "Age (ka)",
               ylab = "Frequency")$breaks

# Calculate the observed frequencies
observed <- hist(age_acrop, breaks = breaks, plot = FALSE)$counts

# Monte Carlo Simulation
n_sim <- 10000
simulated_gaps <- numeric(n_sim)
simulated_gaps_size <- vector(mode = "list", length = n_sim)

for (i in 1:n_sim) {
  simulated_data <- runif(81, min = min(age_acrop), max = max(age_acrop))
  
  sim_observed <- hist(simulated_data, breaks = CI, plot = FALSE)$counts
  
  simulated_gaps[i] <- sum(sim_observed == 0)
  simulated_gaps_size[[i]] <- gaps(sim_observed)
}

# Gap size analysis

gap_size <- list_rbind(simulated_gaps_size, names_to = "simulation")|>
  group_by(simulation, size)|>
  reframe(f = n())

barbados <- gap_size|>
  group_by(size)|>
  summarise(average = round(mean(f),0), times = length(size), freq = (times+1)/(n_sim+1))|>
  mutate(site = "Barbados")
```

## Belize

```{r}
acrop <-raw_ages|>
  filter(source == "Gischler et al.,23")|>
  filter(species_2 == "BRAN")

age_acrop <- acrop$t_ka

age_range <- 1000 * (max(age_acrop) - min(age_acrop))

intervals <-100

CI <- round(age_range/intervals, 0)

# Histogram to visualize the data
hist(age_acrop,
     breaks = CI, 
     main = paste("Barbados,", intervals, "years intervals"),
     xlab = "Age (ka)",
     ylab = "Frequency")

# Defining class intervals
breaks <- hist(age_acrop, 
               breaks = CI, 
               main = paste("Barbados,", intervals, "years intervals"),
               xlab = "Age (ka)",
               ylab = "Frequency")$breaks

# Calculate the observed frequencies
observed <- hist(age_acrop, breaks = breaks, plot = FALSE)$counts

# Monte Carlo simulation
n_sim <- 10000
simulated_gaps <- numeric(n_sim)
simulated_gaps_size <- vector(mode = "list", length = n_sim)

for (i in 1:n_sim) {
  simulated_data <- runif(69, min = min(age_acrop), max = max(age_acrop))
  
  sim_observed <- hist(simulated_data, breaks = CI, plot = FALSE)$counts
  
  simulated_gaps[i] <- sum(sim_observed == 0)
  simulated_gaps_size[[i]] <- gaps(sim_observed)
}

# Gap size analysis

gap_size <- list_rbind(simulated_gaps_size, names_to = "simulation")|>
  group_by(simulation, size)|>
  reframe(f = n())

belize <- gap_size|>
  group_by(size)|>
  summarise(average = round(mean(f),0), times = length(size), freq = (times+1)/(n_sim+1))|>
  mutate(site = "Belize")
```

## St. Croix

```{r}
acrop <-raw_ages|>
  filter(source == "Hubbard et al.,05;13")|>
  filter(species_2 == "BRAN")

age_acrop <- acrop$t_ka

age_range <- 1000 * (max(age_acrop) - min(age_acrop))

intervals <- 100

CI <- round(age_range/intervals, 0)

# Histogram to visualize the data
hist(age_acrop,
     breaks = CI, 
     main = paste("Saint Croix,", intervals, "years intervals"),
     xlab = "Age (ka)",
     ylab = "Frequency")

# Defining class intervals
breaks <- hist(age_acrop, 
               breaks = CI, 
               main = paste("Saint Croix,", intervals, "years intervals"),
               xlab = "Age (ka)",
               ylab = "Frequency")$breaks

# Calculate the observed frequencies
observed <- hist(age_acrop, breaks = breaks, plot = FALSE)$counts

# Monte Carlo simulations
n_sim <- 10000
simulated_gaps <- numeric(n_sim)
simulated_gaps_size <- vector(mode = "list", length = n_sim)

for (i in 1:n_sim) {
  simulated_data <- runif(52, min = min(age_acrop), max = max(age_acrop))
  
  sim_observed <- hist(simulated_data, breaks = CI, plot = FALSE)$counts
  
  simulated_gaps[i] <- sum(sim_observed == 0)
  simulated_gaps_size[[i]] <- gaps(sim_observed)
}

# Gap size analysis
gap_size <- list_rbind(simulated_gaps_size, names_to = "simulation")|>
  group_by(simulation, size)|>
  reframe(f = n())

stc <- gap_size|>
  group_by(size)|>
  summarise(average = round(mean(f),0), times = length(size), freq = (times+1)/(n_sim+1))|>
  mutate(site = "Saint Croix")
```

### Regional scope for large gap probabilities

```{r}
raw_ages <- read_excel("raw_ages.xlsx")
acrop <-raw_ages|>
  filter(species_2 == "BRAN")

age_acrop <- acrop$t_ka

age_range <- 1000 * (max(age_acrop) - min(age_acrop))

intervals <- 100

CI <- round(age_range/intervals, 0)

# Histogram to visualize the data
hist(age_acrop,
     breaks = CI, 
     main = paste("All sites,", intervals, "years intervals"),
     xlab = "Age (ka)",
     ylab = "Frequency")

# Defining class intervals
breaks <- hist(age_acrop, 
               breaks = CI, 
               main = paste("All sites,", intervals, "years intervals"),
               xlab = "Age (ka)",
               ylab = "Frequency")$breaks

# Calculate the observed frequencies
observed <- hist(age_acrop, breaks = breaks, plot = FALSE)$counts

# Monte Carlo simulation
n_sim <- 10000
simulated_gaps <- numeric(n_sim)
simulated_gaps_size <- vector(mode = "list", length = n_sim)

for (i in 1:n_sim) {
  simulated_data <- runif(266, min = min(age_acrop), max = max(age_acrop))
  
  sim_observed <- hist(simulated_data, breaks = CI, plot = FALSE)$counts
  
  simulated_gaps[i] <- sum(sim_observed == 0)
  simulated_gaps_size[[i]] <- gaps(sim_observed)
}

# Gap size analysis

gap_size <- list_rbind(simulated_gaps_size, names_to = "simulation")|>
  group_by(simulation, size)|>
  reframe(f = n())

all_sites <- gap_size|>
  group_by(size)|>
  summarise(average = round(mean(f),0), times = length(size), freq = (times+1)/(n_sim+1))|>
  mutate(site = "All sites")


bind_rows(yucatan, barbados, belize, stc, all_sites) |>
  filter(size <= 1300) |>
  mutate(Region = factor(
    site,
    levels = c("Yucatan", "Barbados", "Belize", "Saint Croix", "All sites"),
    labels = c(
      "Yucatan (n = 64)",
      "Barbados (n = 81)",
      "Belize (n = 69)",
      "Saint Croix (n = 52)",
      "All sites (n = 266)"
    )
  )) |>
  ggplot(aes(x = size, y = freq)) +
  scale_y_continuous(breaks = seq(0, 1, 0.2)) +
  scale_x_continuous(breaks = seq(50, 14000, 100)) +
  geom_hline(yintercept = 0.05,
             colour = "black",
             linetype="longdash",
             linewidth = 1) +
  ylab(expression("Probability under " * H[0] * " of uniform distribution") ) +
  xlab("Gap size (years)") +
  geom_line(aes(colour = Region), linewidth = 1) +
  theme_bw() +
  theme(
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank(),
    panel.grid.minor.y = element_blank(),
    panel.grid.major.y = element_blank(),
    legend.position = "inside",
    legend.justification = c(0.95, 0.95),
    legend.background = element_rect(fill = "white", colour = "white")
  )

```


